---
title: "SEC 1-FA8 - DELA ROSA, R"
author: "Roland Dela Rosa"
date: "May 1, 2025"
output:
  pdf_document
subtitle: "**GitHub Link:(https://github.com/rddelarosa/APM1110/blob/main/FA8/SEC_1-FA8-DELA-ROSA%2C-R.md?plain=1)**"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. An analogue signal received at a detector, measured in microvolts, is normally distributed with a mean of 200 and variance of 256.

### Given variables:
```{r}
mu <- 200
vu <- sqrt(256)
```

(a) What is the probability that the signal will exceed 224 $\mu$V?
```{r}
p_exceed <- 1 - pnorm(224, mu, vu)
p_exceed
```

(b) What is the probability that it will be between 186 and 224 $\mu$V?
```{r}
p_between <- pnorm(224, mu, vu) - pnorm(186, mu, vu)
p_between
```

(c) What is the micro voltage below which 25% of the signals will be?
```{r}
v25_percent <- qnorm(0.25, mu, vu)
v25_percent
```

(d) What is the probability that the signal will be less than 240 $\mu$V, given that it is larger than 210 $\mu$V?
```{r}
numerator_240 <- pnorm(240, mu, vu) - pnorm(210, mu, vu)
denominator_240 <- 1 - pnorm(210, mu, vu)
cond_prob_240 <- numerator_240 / denominator_240
cond_prob_240
```

(e) Estimate the interquartile range.
```{r}
Q1 <- qnorm(0.25, mu, vu)
Q3 <- qnorm(0.75, mu, vu)
IQR <- Q3 - Q1
IQR
```


(f) What is the probability that the signal will be less than 220 $\mu$V, given that it is larger than 210 $\mu$V?
```{r}
numerator_220 <- pnorm(220, mu, vu) - pnorm(210, mu, vu)
denominator_220 <- 1 - pnorm(210, mu, vu)
cond_prob_220 <- numerator_220 / denominator_220
cond_prob_220
```

(g) If we know that a received signal is greater than 200 $\mu$V, what is the probability that it is in fact greater than 220 $\mu$V?
```{r}
numerator_200 <- 1 - pnorm(220, mu, vu)
denominator_200 <- 1 - pnorm(200, mu, vu)
cond_prob_200 <- numerator_200 / denominator_200
cond_prob_200
```

## 2. A manufacturer of a particular type of computer system is interested in improving its customer support services. Over a period of six months, customers were surveyed and the amount of downtime (in minutes) due to system failures was recorded.

### Given variables
```{r}
md <- 25
vd <- sqrt(144)
```

(a) Obtain bounds which will include 95% of the downtime of all the customers
```{r}
lower_bound <- qnorm(0.025, md, vd)
upper_bound <- qnorm(0.975, md, vd)
cat("The lower bound is",lower_bound,"\nThe upper bound is", upper_bound)

```

(b) Obtain the bound above which 10% of the downtime is included

```{r}
downtime_90th_percentile <- qnorm(0.90, md, vd)
downtime_90th_percentile
```
