---
title: "SEC_1-FA1_GROUP_6-DELA ROSA, R"
subtitle: "Github Link: https://github.com/rddelarosa/APM1110/tree/ac696deeb7caea0bf384eca2a8a9d8cc6041a182/FA1"
author: "Roland Dela Rosa"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Write the skewness program, and use it to calculate the skewness coefficient of the four examination subjects in results.txt (results.csv). What can you say about these data?

## Pearson has given an approximate formula for the skewness that is easier to calculate than the exact formula given in Equation 2.1.

### Read the dataset
```{r}
data <- read.table("results.txt", header = TRUE)
```

### Compute Normal Skewness
```{r}
skew_n <- function(x) {
  mean_x <- mean(x, na.rm = TRUE)  
  sum1 <- sum((x-mean_x) ** 2, na.rm = TRUE)
  sum2 <- sum((x-mean_x) ** 3, na.rm = TRUE)
  skew <- (sqrt(length(x))* sum2)/(sum1 **(3/2))
  skew
}
subjects <- c("Arch1", "Prog1", "Arch2", "Prog2")
skewness_exact <- sapply(subjects, function(subj) skew_n(data[[subj]]))
print(skewness_exact)
```
### Compute Pearson's skewness for each subject
```{r}
skew_p <- function(x) {
  mean_x <- mean(x, na.rm = TRUE)
  median_x <- median(x, na.rm = TRUE)
  sd_x <- sd(x, na.rm = TRUE)
  
  return(3 * (mean_x - median_x) / sd_x)
}

skewness_values <- sapply(subjects, function(subj) skew_p(data[[subj]]))
print(skewness_values)
```


### Boxplot for Skewness Visualization
```{r}
boxplot(data[, subjects], main = "Boxplot of Skewness Distribution", col = c("red", "green", "blue", "purple"), 
        names = subjects, ylab = "Scores")
```

### Boxplot for Pearson's and Normal Skewness
```{r}
boxplot(skewness_values, skewness_exact, names = c("Pearson", "Normal"), col = c("orange", "blue"),
        main = "Comparison of Pearson and Normal Skewness", ylab = "Skewness Coefficient")
```

### Interpretation
#### Pearson's Skewness
Pearson's skewness is a simpler measure that estimates skewness using mean, median, and standard deviation. It is useful when a quick approximation is needed but may not always capture extreme skewness accurately.

#### Normal Skewness
The normal skewness calculation accounts for the third moment of the data distribution, providing a more precise measure of asymmetry. This method is generally more accurate for highly skewed distributions.

#### Comparison
Comparing both values, Pearson's skewness is generally close to the normal skewness but may deviate when the data distribution is highly skewed or has extreme outliers. The boxplot comparing both skewness values visually represents their differences across different subjects.

---

## 2.  For the class of 50 students of computing detailed in Exercise 1.1, use R to 

## (a)  form the stem-and-leaf display for each gender, and discuss the advantages of this representation compared to the traditional histogram; 

## (b)  construct a box-plot for each gender and discuss the findings. 

### Define the data
```{r}
females <- c(57, 59, 78, 79, 60, 65, 68, 71, 75, 48, 51, 55, 56, 41, 43, 
             44, 75, 78, 80, 81, 83, 83, 85)
males <- c(48, 49, 49, 30, 30, 31, 32, 35, 37, 41, 86, 42, 51, 53, 56,
           42, 44, 50, 51, 65, 67, 51, 56, 58, 64, 64, 75)
```
### (a) Stem-and-Leaf Display
```{r}
cat("\nStem-and-Leaf for Females:\n")
stem(females)

cat("\nStem-and-Leaf for Males:\n")
stem(males)
```
### Interpretation of Stem-and-Leaf vs. Histogram
The stem-and-leaf plot preserves individual data points while showing the overall shape of the distribution. Compared to a histogram, it provides a more detailed view of the actual values while still displaying the distribution.

---

### (b) Boxplot Comparison
```{r}
boxplot(females, males, names = c("Females", "Males"), col = c("pink", "blue"), 
        main = "Boxplot of Scores by Gender", ylab = "Scores")
```

### Interpretation
The boxplot allows us to visualize the spread, median, and presence of outliers. If the median line is not centered, it indicates skewness. The interquartile range (IQR) shows score variability, and outliers (if present) highlight extreme values. 

From the boxplot, we can compare the central tendency and dispersion between male and female students.

---

